Analyzing Strawberry Pricing Strategies Across Canadian Vendors

### 2024-12-03 paper/paper.qmd
render the paper

- **Vendor (`vendor`)**: A categorical variable identifying the supplier of strawberries. Different vendors adopt distinct pricing strategies that influence strawberry prices.

- **Original Price (`old_price`)**: A continuous variable representing the original price of strawberries before discounts or promotions. This variable reflects vendor-specific baseline pricing strategies.

- **Month (`month`)**: A numeric variable capturing the observation month, enabling the analysis of seasonal trends and time-based fluctuations.

- **Price Gap (`old_price - current_price`)**: A derived variable quantifying the magnitude of discounts or markdowns applied by vendors.

**Model set-up**

@@

- \( y_i \): Current price (`current_price`) of strawberries for observation \( i \) (dependent variable).
- \( \beta_0 \): Intercept, representing the baseline price (reference vendor, average original price, and earliest month).
- \( \beta_{\text{vendor}[i]} \): Vendor-specific effect, capturing how each supplier’s prices deviate from the baseline vendor.
- \( \beta_{\text{old\_price}} \): Effect of the original price, representing how the initial pricing influences the current price.
- \( \beta_{\text{month}} \): Effect of the observation month, capturing seasonal or temporal trends in strawberry pricing.
- \( \epsilon_i \): Residual error, accounting for variability not explained by the predictors.

@@

Coefficients (\( \beta \)):

@@

Residual Standard Deviation (\( \sigma \)):

@@

The priors in this analysis reflect reasonable assumptions about the relationships between predictors and strawberry prices. Coefficients $(\beta)$ are assigned normal priors $(\operatorname{Normal}(0,2.5)$ ) to reflect the expectation that most effects are small and centered around zero, with moderate variability. For example, $\beta_{\text {old_price }}$ is expected to be positive, reflecting the influence of original prices on current prices, while $\beta_{\text {month }}$ accounts for typically small seasonal pricing effects. The intercept $\left(\beta_0\right)$ also follows a Normal $(0,2.5)$, assuming the baseline price is near zero with moderate deviations. Finally, the residual standard deviation $(\sigma)$ is given an exponential prior $(\operatorname{Exponential}(1))$ to ensure positivity and account for variability not captured by the predictors.

@@

1. Uncertainty Quantification:

   - The Bayesian approach provides posterior distributions for each parameter, allowing for uncertainty quantification. For example, the posterior distribution of \( \beta_{\text{vendor}} \) helps quantify the uncertainty around vendor-specific pricing effects.

2. Incorporating Prior Knowledge:

   - Priors enable the incorporation of domain knowledge into the analysis. For instance:
     - \( \beta_{\text{vendor}} \): Reflects the expectation that most vendor effects are close to zero, with some deviations.
     - \( \beta_{\text{month}} \): Captures small but meaningful seasonal effects.
     - \( \beta_{\text{old\_price}} \): Represents the assumption that higher original prices lead to higher observed prices.

3. Handling Sparse Data:

@@

4. Probabilistic Insights:

   - Posterior distributions provide credible intervals, making it easier to interpret parameter estimates. For instance, a 95% credible interval for \( \beta_{\text{old\_price}} \) of \([0.8, 1.2]\) means there is a 95% probability that the original price effect lies within this range.

@@

modelsummary::modelsummary(
  list(
    "strawberry price model" = vendor_price_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

## Comprehensive Data Collection Plan for Strawberry Pricing Analysis

@@

Each coefficient (\(\beta\)) will have a prior distribution, such as a normal prior centered at zero, reflecting initial uncertainty about its effect. For instance, the prior for \(\beta_{\text{vendor}}\) assumes no significant differences between vendors initially, allowing the data to dominate posterior estimates. The Bayesian framework also provides posterior distributions for each parameter, which capture uncertainty and enable credible intervals for inference. For example, a 95% credible interval for \(\beta_{\text{old\_price}}\) indicates the range of plausible values for the effect of original prices on current prices, accounting for uncertainty.

@@

The primary dependent variable in the model is `current_price`, representing the observed price per pound of strawberries. Key predictors include vendor identity (`vendor`), original price (`old_price`), sale status (`sale_status`), and temporal trends (`month`). These variables capture the diverse factors influencing pricing, such as competitive strategies among vendors, promotional effects, and seasonal supply-demand dynamics.

The model specification includes normal priors for regression coefficients (\(\beta\)) and an exponential prior for the residual standard deviation (\(\sigma\)). This setup reflects reasonable assumptions about the effects of predictors, such as small, centered effects with moderate variability. For instance, the prior for \(\beta_{\text{vendor}}\) assumes no extreme pricing differences between vendors, while allowing the data to dominate posterior updates.

>>>>

- **Vendor ($\text{vendor}$)**: A categorical variable identifying the supplier of strawberries. Different vendors adopt distinct pricing strategies that influence strawberry prices.

- **Original Price ($\text{old\_price}$)**: A continuous variable representing the original price of strawberries before discounts or promotions. This variable reflects vendor-specific baseline pricing strategies.

- **Month ($\text{month}$)**: A numeric variable capturing the observation month, enabling the analysis of seasonal trends and time-based fluctuations.

- **Price Gap ($\text{old\_price} - \text{current\_price}$)**: A derived variable quantifying the magnitude of discounts or markdowns applied by vendors.

**Model Set-Up**

@@

- $y_i$: Current price ($\text{current\_price}$) of strawberries for observation $i$ (dependent variable).  
- $\beta_0$: Intercept, representing the baseline price (reference vendor, average original price, and earliest month).  
- $\beta_{\text{vendor}[i]}$: Vendor-specific effect, capturing how each supplier’s prices deviate from the baseline vendor.  
- $\beta_{\text{old\_price}}$: Effect of the original price, representing how the initial pricing influences the current price.  
- $\beta_{\text{month}}$: Effect of the observation month, capturing seasonal or temporal trends in strawberry pricing.  
- $\epsilon_i$: Residual error, accounting for variability not explained by the predictors.

@@

For the coefficients ($\beta$):

@@

For the residual standard deviation ($\sigma$):

@@

The priors in this analysis reflect reasonable assumptions about the relationships between predictors and strawberry prices. Coefficients ($\beta$) are assigned normal priors ($\text{Normal}(0, 2.5)$) to reflect the expectation that most effects are small and centered around zero, with moderate variability. For example, $\beta_{\text{old\_price}}$ is expected to be positive, reflecting the influence of original prices on current prices, while $\beta_{\text{month}}$ accounts for typically small seasonal pricing effects. The intercept ($\beta_0$) also follows a $\text{Normal}(0, 2.5)$, assuming the baseline price is near zero with moderate deviations. Finally, the residual standard deviation ($\sigma$) is given an exponential prior ($\text{Exponential}(1)$) to ensure positivity and account for variability not captured by the predictors.

@@

1. **Uncertainty Quantification**:

   - The Bayesian approach provides posterior distributions for each parameter, allowing for uncertainty quantification. For example, the posterior distribution of $\beta_{\text{vendor}}$ helps quantify the uncertainty around vendor-specific pricing effects.

2. **Incorporating Prior Knowledge**:

   - Priors enable the incorporation of domain knowledge into the analysis. For instance:
     - $\beta_{\text{vendor}}$: Reflects the expectation that most vendor effects are close to zero, with some deviations.
     - $\beta_{\text{month}}$: Captures small but meaningful seasonal effects.
     - $\beta_{\text{old\_price}}$: Represents the assumption that higher original prices lead to higher observed prices.

3. **Handling Sparse Data**:

@@

4. **Probabilistic Insights**:

   - Posterior distributions provide credible intervals, making it easier to interpret parameter estimates. For instance, a 95% credible interval for $\beta_{\text{old\_price}}$ of $[0.8, 1.2]$ means there is a 95% probability that the original price effect lies within this range.

@@

#modelsummary::modelsummary(
  #list(
  #  "strawberry price model" = vendor_price_model
  #),
  #statistic = "mad",
  #fmt = 2
#)
```

@@

# Comprehensive Data Collection Plan for Strawberry Pricing Analysis

@@

Each coefficient ($\beta$) will have a prior distribution, such as a normal prior centered at zero, reflecting initial uncertainty about its effect. For instance, the prior for $\beta_{\text{vendor}}$ assumes no significant differences between vendors initially, allowing the data to dominate posterior estimates. The Bayesian framework also provides posterior distributions for each parameter, which capture uncertainty and enable credible intervals for inference. For example, a 95% credible interval for $\beta_{\text{old\_price}}$ indicates the range of plausible values for the effect of original prices on current prices, accounting for uncertainty.

@@

The primary dependent variable in the model is `current_price`, representing the observed price per pound of strawberries. Key predictors include vendor identity ($\text{vendor}$), original price ($\text{old\_price}$), sale status ($\text{sale\_status}$), and temporal trends ($\text{month}$). These variables capture the diverse factors influencing pricing, such as competitive strategies among vendors, promotional effects, and seasonal supply-demand dynamics.

The model specification includes normal priors for regression coefficients ($\beta$) and an exponential prior for the residual standard deviation ($\sigma$). This setup reflects reasonable assumptions about the effects of predictors, such as small, centered effects with moderate variability. For instance, the prior for $\beta_{\text{vendor}}$ assumes no extreme pricing differences between vendors, while allowing the data to dominate posterior updates.
